---
title: "Percent of Normal Maps"
author: "Will Carrara"
date: "4/21/2020"
output: 
  bookdown::html_document2
---

<style>
.main-container {max-width: 1200px !important;}
  body {text-align: justify}
</style>

<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"https://www.nasa.gov/sites/default/files/thumbnails/image/nasa-logo-web-rgb.png\" style=\"float: right;width: 200px;\"/>')});
</script>

***

### Abstract{-}

This tool is used to generate maps which explore the phenology of individual crops over a predefined 5 year period for the state of California. The historical “non-drought" years used as reference are 2008, 2009, 2010, 2013, 2017. Each colored pixel represents the average NDVI value for a crop partitioned by season as a percent of the same period for the reference years.

```{r include=FALSE}
knitr::opts_chunk$set(comment=NA, echo=FALSE, fig.align="center", warning=FALSE)

library(ggplot2)
theme_set(theme_bw())

library(maps)
library(gridExtra)

library(rgdal)
```

```{r, cache=TRUE}
ca_spring_2018 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Spring_2018.csv")
ca_summer_2018 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Summer_2018.csv")

# lat-lon for each id 
ca_gps = read.csv("spacial/ca_gps.csv")
```


```{r, eval=TRUE}
ca_spring_2015 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Spring_2015.csv")
ca_summer_2015 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Summer_2015.csv")

ca_spring_2016 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Spring_2016.csv")
ca_summer_2016 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Summer_2016.csv")

ca_spring_2017 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Spring_2017.csv")
ca_summer_2017 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Summer_2017.csv")

ca_spring_2018 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Spring_2018.csv")
ca_summer_2018 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Summer_2018.csv")

ca_spring_2019 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Spring_2019.csv")
ca_summer_2019 = read.csv("/Users/willcarrara/Desktop/MISC/FAM_backup/FAM/California/output/California_Summer_2019.csv")

# lat-lon for each id 
ca_gps = read.csv("spacial/ca_gps.csv")
```


```{r, eval=FALSE}
# data from the FAM outputs
ca_spring_2015 = read.csv("../output/California_Spring_2015.csv")
ca_summer_2015 = read.csv("../output/California_Summer_2015.csv")

ca_spring_2016 = read.csv("../output/California_Spring_2016.csv")
ca_summer_2016 = read.csv("../output/California_Summer_2016.csv")

ca_spring_2017 = read.csv("../output/California_Spring_2017.csv")
ca_summer_2017 = read.csv("../output/California_Summer_2017.csv")

ca_spring_2018 = read.csv("../output/California_Spring_2018.csv")
ca_summer_2018 = read.csv("../output/California_Summer_2018.csv")

ca_spring_2019 = read.csv("../output/California_Spring_2019.csv")
ca_summer_2019 = read.csv("../output/California_Summer_2019.csv")

# lat-lon for each id 
ca_gps = read.csv("spacial/ca_gps.csv")
```


```{r, cache=TRUE}
process = function(fam_output) {
  # add lat lon 
  fam_output$lat = ca_gps$lat
  fam_output$lon = ca_gps$lon
  
  # take only postive values
  fam_output = fam_output[fam_output$percent_5yr_Avg >= 0 & fam_output$percent_5yr_Avg < 400,] 
  
  # add dummy group column for plotting
  fam_output$group = -1
  
  # add breaks
  fam_output$bin = cut(fam_output$percent_5yr_Avg, breaks = c(0,60,80,max(fam_output$percent_5yr_Avg)), labels=c("0-60", "60-80","80-max"))

  return(fam_output)
}

# create new dataframe 
ca_spring_2015 = process(ca_spring_2015)
ca_summer_2015 = process(ca_summer_2015)

ca_spring_2016 = process(ca_spring_2016)
ca_summer_2016 = process(ca_summer_2016)

ca_spring_2017 = process(ca_spring_2017)
ca_summer_2017 = process(ca_summer_2017)

ca_spring_2018 = process(ca_spring_2018)
ca_summer_2018 = process(ca_summer_2018)

ca_spring_2019 = process(ca_spring_2019)
ca_summer_2019 = process(ca_summer_2019)
```


```{r}
# uses map package to get state information# usa
states <- map_data("state")

# california
ca_df <- states[states$region== "california",]
counties <- map_data("county")
ca_county <- counties[counties$region== "california",]
```


```{r}
# theme options
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank(),
  plot.title = element_text(face="bold", hjust=0.18, vjust=-3, size=18),
  legend.position=c(.58, 0.81)
)

# plotting function
plotter= function(fam_output_spring, fam_output_summer, state_df, county_df, year) {
  # spring season
  ca_spring <- ggplot(data=state_df, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Spring", year)) +
  geom_polygon(color="black", fill="gray") + geom_polygon(data=county_df, fill=NA, color="white") +
  geom_point(data=fam_output_spring, aes(x=lon, y=lat, colour=percent_5yr_Avg), shape=".", show.legend=TRUE) +
  geom_polygon(color="black", fill=NA) + ditch_the_axes + 
  scale_color_gradient2(low="orange4", high="forestgreen", mid="khaki", midpoint=50, name="Percent Normal", breaks=c(.5,25,50,75,100), labels=c("0","25","50","75","100"), limits=c(0, 115), na.value = "chartreuse4")
  
  #ggsave(paste0("Spring_", year, ".png"), dpi=1200)
  
  # summer season
  ca_summer <- ggplot(data=state_df, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Summer", year)) +
  geom_polygon(color="black", fill="gray") + geom_polygon(data=county_df, fill=NA, color="white") +
  geom_point(data=fam_output_summer, aes(x=lon, y=lat, colour=percent_5yr_Avg), shape=".", show.legend=TRUE) +
  geom_polygon(color="black", fill=NA) + ditch_the_axes + 
  scale_color_gradient2(low="orange4", high="forestgreen", mid="khaki", midpoint=50, name="Percent Normal", breaks=c(.5,25,50,75,100), labels=c("0","25","50","75","100"), limits=c(0, 115), na.value = "chartreuse4")
  
  #ggsave(paste0("Summer_", year, ".png"), dpi=1200)
  
  # side by side plot
  grid.arrange(ca_spring, ca_summer, ncol=2)

}
```


```{r fig.width=18, fig.asp=.5}
# 2015
#plotter(ca_spring_2015, ca_summer_2015, ca_df, ca_county, "2015")

# 2016
#plotter(ca_spring_2016, ca_summer_2016, ca_df, ca_county, "2016")

# 2017
#plotter(ca_spring_2017, ca_summer_2017, ca_df, ca_county, "2017")

# 2018
#plotter(ca_spring_2018, ca_summer_2018, ca_df, ca_county, "2018")

# 2019
#plotter(ca_spring_2019, ca_summer_2019, ca_df, ca_county, "2019")
```



```{r, fig.width=18, fig.asp=.5}
path = "/Users/willcarrara/Downloads/cb_2018_06_sldl_500k/cb_2018_06_sldl_500k.shp"
shp <- readOGR(path)
ca_shp = spTransform(shp, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))

# plotting function
plotter2= function(fam_output_spring, fam_output_summer, year) {
  
   # spring season
   ca_spring <- ggplot(data=ca_shp, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Spring", year)) +
   geom_polygon(color="gray25", fill=alpha("#2E4053",.65)) + ditch_the_axes + 
   geom_point(data=fam_output_spring, aes(x=lon, y=lat, colour=bin), shape=".", show.legend=TRUE) +
   scale_colour_manual(values=c("orange4", "khaki", "forestgreen"), na.translate = F) + 
  guides(colour = guide_legend(override.aes=list(shape=15, size=7), title="Percent Normal"))
    
  
  #ggsave(paste0("Spring_", year, ".png"), dpi=1200)
  
  # summer season
   ca_summer <- ggplot(data=ca_shp, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Summer", year)) +
   geom_polygon(color="gray25", fill=alpha("#2E4053",.65)) + ditch_the_axes + 
   geom_point(data=fam_output_summer, aes(x=lon, y=lat, colour=percent_5yr_Avg), shape=".", show.legend=TRUE) +
   scale_color_gradient2(low="orange4", high="forestgreen", mid="khaki", midpoint=50, name="Percent Normal", breaks=c(.5,25,50,75,100), labels=c("0","25","50","75","100"), limits=c(0, 115), na.value="chartreuse4")
  
  #ggsave(paste0("Summer_", year, ".png"), dpi=1200)
  
  # side by side plot
  grid.arrange(ca_spring, ca_summer, ncol=2)
}

# 2015
#plotter2(ca_spring_2015, ca_summer_2015, "2015")

# 2016
#plotter2(ca_spring_2016, ca_summer_2016, "2016")

# 2017
#plotter2(ca_spring_2017, ca_summer_2017, "2017")

# 2018
#plotter2(ca_spring_2018, ca_summer_2018, "2018")

# 2019
#plotter2(ca_spring_2019, ca_summer_2019, "2019")
```

```{r, fig.width=18, fig.asp=.5}
path = "/Users/willcarrara/Downloads/cb_2018_06_sldl_500k/cb_2018_06_sldl_500k.shp"
shp <- readOGR(path)
ca_shp = spTransform(shp, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))

low =  alpha("orange4",.7)
mid =  alpha("khaki",.7)
high =  alpha("forestgreen",.7)

# plotting function
plotter3 = function(fam_output_spring, fam_output_summer, year) {
  
   # spring season
   ca_spring <- ggplot(data=ca_shp, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Spring", year)) +
   geom_polygon(color="gray25", fill=alpha("#2E4053",.65)) + ditch_the_axes + 
   geom_point(data=fam_output_spring, aes(x=lon, y=lat, colour=bin), shape=".", show.legend=TRUE) +
   scale_colour_manual(values=c(low, mid, high), na.translate = F) + 
  guides(colour = guide_legend(override.aes=list(shape=15, size=7), title="Percent Normal"))
    
  
  #ggsave(paste0("Spring_", year, ".png"), dpi=1200)
  
  # summer season
   ca_summer <- ggplot(data=ca_shp, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Summer", year)) +
   geom_polygon(color="gray25", fill=alpha("#2E4053",.65)) + ditch_the_axes + 
   geom_point(data=fam_output_summer, aes(x=lon, y=lat, colour=bin), shape=".", show.legend=TRUE) +
   scale_colour_manual(values=c("orange4", "khaki", "forestgreen"), na.translate = F) + 
  guides(colour = guide_legend(override.aes=list(shape=15, size=7), title="Percent Normal"))
  
  #ggsave(paste0("Summer_", year, ".png"), dpi=1200)
  
  # side by side plot
  grid.arrange(ca_spring, ca_summer, ncol=2)
}

# 2015
plotter3(ca_spring_2015, ca_summer_2015, "2015")

# 2016
plotter3(ca_spring_2016, ca_summer_2016, "2016")

# 2017
plotter3(ca_spring_2017, ca_summer_2017, "2017")

# 2018
plotter3(ca_spring_2018, ca_summer_2018, "2018")

# 2019
plotter3(ca_spring_2019, ca_summer_2019, "2019")
```
