---
title: "Percent of Normal Maps"
author: "Will Carrara"
date: "4/21/2020"
output: 
  bookdown::html_document2
---

<style>
.main-container {max-width: 1200px !important;}
  body {text-align: justify}
</style>

<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"https://www.nasa.gov/sites/default/files/thumbnails/image/nasa-logo-web-rgb.png\" style=\"float: right;width: 200px;\"/>')});
</script>

***

### Abstract{-}

This tool is used to generate maps which explore the phenology of individual crops over a predefined 5 year period for the state of California. The historical “non-drought" years used as reference are 2008, 2009, 2010, 2013, 2017. Each colored pixel represents the average NDVI value for a crop partitioned by season as a percent of the same period for the reference years.

```{r include=FALSE}
knitr::opts_chunk$set(comment=NA, echo=FALSE, fig.align="center", warning=FALSE)

library(ggplot2)
theme_set(theme_bw())

library(maps)
library(gridExtra)
```


```{r}
# data from the FAM outputs
ca_spring_2016 = read.csv("../output/California_Spring_2016.csv")
ca_summer_2016 = read.csv("../output/California_Summer_2016.csv")

ca_spring_2017 = read.csv("../output/California_Spring_2017.csv")
ca_summer_2017 = read.csv("../output/California_Summer_2017.csv")

ca_spring_2018 = read.csv("../output/California_Spring_2018.csv")
ca_summer_2018 = read.csv("../output/California_Summer_2018.csv")

ca_spring_2019 = read.csv("../output/California_Spring_2019.csv")
ca_summer_2019 = read.csv("../output/California_Summer_2019.csv")

# lat-lon for each id 
ca_gps = read.csv("spacial/ca_gps.csv")
```


```{r}
process = function(fam_output) {
  # add lat lon 
  fam_output$lat = ca_gps$lat
  fam_output$lon = ca_gps$lon
  
  # take only postive values
  fam_output = fam_output[fam_output$percent_5yr_Avg >= 0 & fam_output$percent_5yr_Avg <= 350,] 
  
  # create bins
  fam_output$bin  <- cut(fam_output$percent_5yr_Avg, breaks=c(0,20,40,60,80,100,120), labels=c("20","40","60","80","100","120"))
  fam_output$bin = as.numeric(as.character(fam_output$bin))
  
  # add dummy group column for plotting
  fam_output$group = 7 
  
  return(fam_output)
}

# create new dataframe 
ca_spring_2016 = process(ca_spring_2016)
ca_summer_2016 = process(ca_summer_2016)

ca_spring_2017 = process(ca_spring_2017)
ca_summer_2017 = process(ca_summer_2017)

ca_spring_2018 = process(ca_spring_2018)
ca_summer_2018 = process(ca_summer_2018)

ca_spring_2019 = process(ca_spring_2019)
ca_summer_2019 = process(ca_summer_2019)
```


```{r}
# uses map package to get state information# usa
states <- map_data("state")

# california
ca_df <- states[states$region== "california",]
counties <- map_data("county")
ca_county <- counties[counties$region== "california",]
```


```{r}
# theme options
ditch_the_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank(),
  plot.title = element_text(face="bold", hjust=0.18, vjust=-3, size=18),
  legend.position=c(.58, 0.81)
)

# plotting function
plotter= function(fam_output_spring, fam_output_summer, state_df, county_df, year) {
  # spring season
  ca_spring <- ggplot(data=state_df, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Spring", year)) +
  geom_polygon(color="black", fill="gray") + geom_polygon(data=county_df, fill=NA, color="white") +
  geom_point(data=fam_output_spring, aes(x=lon, y=lat, colour=bin), shape = ".", show.legend=TRUE) +
  geom_polygon(color="black", fill=NA) + ditch_the_axes + 
  scale_color_gradient2(low="orange4", high="forestgreen", midpoint=75, mid="khaki", name="% normal")
  
  # summer season
  ca_summer <- ggplot(data=state_df, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Summer", year)) +
  geom_polygon(color="black", fill="gray") + geom_polygon(data=county_df, fill=NA, color="white") +
  geom_point(data=fam_output_summer, aes(x=lon, y=lat, colour=bin), shape = ".", show.legend=TRUE) +
  geom_polygon(color="black", fill=NA) + ditch_the_axes + 
  scale_color_gradient2(low="orange4", high="forestgreen", midpoint=75, mid="khaki", name="% normal")
  
  # side by side plot
  grid.arrange(ca_spring, ca_summer, ncol=2)
}
```


```{r fig.width=18, fig.asp=.5}
# 2016
plotter(ca_spring_2016, ca_summer_2016, ca_df, ca_county, "2016")

# 2017
plotter(ca_spring_2017, ca_summer_2017, ca_df, ca_county, "2017")

# 2018
plotter(ca_spring_2018, ca_summer_2018, ca_df, ca_county, "2018")

# 2019
plotter(ca_spring_2019, ca_summer_2019, ca_df, ca_county, "2019")
```