---
title: "Percent of Normal Maps"
author: "Will Carrara"
date: "4/21/2020"
output: 
  bookdown::html_document2
---

<style>
.main-container {max-width: 1200px !important;}
  body {text-align: justify}
</style>

<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"https://www.nasa.gov/sites/default/files/thumbnails/image/nasa-logo-web-rgb.png\" style=\"float: right;width: 200px;\"/>')});
</script>

***

### Abstract{-}

This tool is used to generate maps which explore the phenology of individual crops over a predefined 5 year period for the the Western United States. The historical “non-drought" years used as reference are 2008, 2009, 2010, 2013, 2017. Each colored pixel represents the average NDVI value for a crop partitioned by season as a percent of the same period for the combined reference years.

```{r, include=FALSE}
# global settings 
knitr::opts_chunk$set(comment=NA, echo=FALSE, fig.align="center", warning=FALSE)

# required libraries 
library(maps)
library(rgdal)
library(ggplot2)
library(gridExtra)

# plot theme settings
theme_set(theme_bw())

# color palette
low = "orange4"
mid = "khaki"
high = "forestgreen"
```

```{r}
# theme options
remove_axes <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank(),
  plot.title = element_text(face="bold", size=18)
)
```

```{r, fig.width=18, fig.asp=.5}
# plotting function California
plotter_ca = function(fam_spring, fam_summer, year, shp) {
    # spring season
    spring <- ggplot(data=shp, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Spring", year)) +
    geom_polygon(color="gray25", fill=alpha("#2E4053",.65)) + remove_axes + 
    geom_point(data=fam_spring, aes(x=lon, y=lat, colour=bin), shape=".", show.legend=TRUE) +
    scale_colour_manual(values=c(low, mid, high), na.translate=F) + 
    guides(colour=guide_legend(override.aes=list(shape=15, size=7), title="Percent Normal")) +
    theme(legend.position=c(.58, 0.81))
    
    #ggsave(paste0("Spring_", year, ".png"), dpi=1200)
    
    # summer season
    summer <- ggplot(data=shp, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Summer", year)) +
    geom_polygon(color="gray25", fill=alpha("#2E4053",.65)) + remove_axes + 
    geom_point(data=fam_summer, aes(x=lon, y=lat, colour=bin), shape=".", show.legend=TRUE) +
    scale_colour_manual(values=c(low, mid, high), na.translate=F) + 
    guides(colour=guide_legend(override.aes=list(shape=15, size=7), title="Percent Normal")) +
    theme(plot.title=element_text(hjust=0.18, vjust=-3, size=18), legend.position=c(.58, 0.81))
    
    #ggsave(paste0("Summer_", year, ".png"), dpi=1200)
  
    # side by side plot
    grid.arrange(spring, summer, ncol=2)
}

# plotting function Washington
plotter_wa = function(fam_spring, fam_summer, year, shp) {
    # spring season
    spring <- ggplot(data=shp, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Spring", year)) +
    geom_polygon(color="gray25", fill=alpha("#2E4053",.65)) + remove_axes + 
    geom_point(data=fam_spring, aes(x=lon, y=lat, colour=bin), shape=".", show.legend=TRUE) +
    scale_colour_manual(values=c(low, mid, high), na.translate=F) +
    guides(colour=guide_legend(override.aes=list(shape=15, size=7), title="Percent Normal")) +
    theme(plot.title=element_text(hjust=0.6, vjust=-2))
    
    #ggsave(paste0("Spring_", year, ".png"), dpi=1200)
    
    # summer season
    summer <- ggplot(data=shp, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Summer", year)) +
    geom_polygon(color="gray25", fill=alpha("#2E4053",.65)) + remove_axes + 
    geom_point(data=fam_summer, aes(x=lon, y=lat, colour=bin), shape=".", show.legend=TRUE) +
    scale_colour_manual(values=c(low, mid, high), na.translate=F) +
    guides(colour=guide_legend(override.aes=list(shape=15, size=7), title="Percent Normal")) +
    theme(plot.title=element_text(hjust=0.6, vjust=-2))
    
    #ggsave(paste0("Summer_", year, ".png"), dpi=1200)
  
    # side by side plot
    grid.arrange(spring, summer, ncol=2)
}

# plotting function Nevada
plotter_nv = function(fam_output, year, shp) {
    # spring season
    year <- ggplot(data=shp, aes(x=long, y=lat, group=group)) + coord_fixed(1.33) + ggtitle(paste("Nevada", year)) +
    geom_polygon(color="gray25", fill=alpha("#2E4053",.65)) + remove_axes + 
    geom_point(data=fam_output, aes(x=lon, y=lat, colour=bin), shape=".", show.legend=TRUE) +
    scale_colour_manual(values=c(low, mid, high), na.translate=F) +
    guides(colour=guide_legend(override.aes=list(shape=15, size=7), title="Percent Normal")) +
      theme(plot.title=element_text(hjust=0.6, vjust=-3))
    
    print(year)
    
    #ggsave(paste0("Spring_", year, ".png"), dpi=1200)
}

```

```{r}
process = function(path, gps) {
    # input data 
    fam_output = read.csv(path)
    
    # intersect
    fam_output = merge(fam_output, gps, by.x="id", by.y="id")
    
    # add lat lon 
    fam_output$lat = gps$lat
    fam_output$lon = gps$lon
    
    # take only postive values
    fam_output = fam_output[fam_output$percent_5yr_Avg >= 0 & fam_output$percent_5yr_Avg < 400,] 
    
    # add dummy group column for plotting
    fam_output$group = -1
    
    # add breaks
    fam_output$bin = cut(fam_output$percent_5yr_Avg, breaks = c(0,60,80,max(fam_output$percent_5yr_Avg)), labels=c("0-60", "60-80","80-max"))
    
    return(fam_output)
}
```


```{r}
state = function(state) {
    if (state == "California") {
        # crop spacial reference
        gps = read.csv("geospacial/ca_gps.csv")
        
        # state base map
        shp = readOGR("geospacial/basemaps/cali/cb_2018_06_sldl_500k.shp", verbose=FALSE)
        base = spTransform(shp, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
        
        # FAM output
        spring_2015 = process("../states/California/output/California_Spring_2015.csv", gps)
        summer_2015 = process("../states/California/output/California_Summer_2015.csv", gps)
        
        spring_2016 = process("../states/California/output/California_Spring_2016.csv", gps)
        summer_2016 = process("../states/California/output/California_Summer_2016.csv", gps)
        
        spring_2017 = process("../states/California/output/California_Spring_2017.csv", gps)
        summer_2017 = process("../states/California/output/California_Summer_2017.csv", gps)
        
        spring_2018 = process("../states/California/output/California_Spring_2018.csv", gps)
        summer_2018 = process("../states/California/output/California_Summer_2018.csv", gps)
        
        spring_2019 = process("../states/California/output/California_Spring_2019.csv", gps)
        summer_2019 = process("../states/California/output/California_Summer_2019.csv", gps)
        
        #2015
        plotter_ca(spring_2015, summer_2015, "2015", shp)
        # 2016
        plotter_ca(spring_2016, summer_2016, "2016", shp)
        # 2017
        plotter_ca(spring_2017, summer_2017, "2017", shp)
        # 2018
        plotter_ca(spring_2018, summer_2018, "2018", shp)
        # 2019
        plotter_ca(spring_2019, summer_2019, "2019", shp)
    }
    
    else if (state == "Nevada") {
        # crop spacial reference
        gps = read.csv("geospacial/nv_gps.csv")
      
        # state base map
        shp = readOGR("geospacial/basemaps/nev/cb_2018_32_sldl_500k.shp", verbose=FALSE)
        base = spTransform(shp, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
        
        # FAM output
        nv_2015 = process("../states/Nevada/output/Nevada_2015.csv", gps)
        nv_2016 = process("../states/Nevada/output/Nevada_2016.csv", gps)
        nv_2017 = process("../states/Nevada/output/Nevada_2017.csv", gps)
        nv_2018 = process("../states/Nevada/output/Nevada_2018.csv", gps)
        nv_2019 = process("../states/Nevada/output/Nevada_2019.csv", gps)

        #2015
        plotter_nv(nv_2015, "2015", shp)
        # 2016
        plotter_nv(nv_2016, "2016", shp)
        # 2017
        plotter_nv(nv_2017, "2017", shp)
        # 2018
        plotter_nv(nv_2018, "2018", shp)
        # 2019
        plotter_nv(nv_2019, "2019", shp)
    }
    
    else if (state == "Washington") {
        # crop spacial reference
        gps = read.csv("geospacial/wa_gps.csv")
        
        # state base map
        shp = readOGR("geospacial/basemaps/wash/cb_2018_53_sldl_500k.shp", verbose=FALSE)
        base = spTransform(shp, CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"))
        
        # FAM output
        spring_2015 = process("../states/Washington/output/Washington_Spring_2015.csv", gps)
        summer_2015 = process("../states/Washington/output/Washington_Summer_2015.csv", gps)
        
        spring_2016 = process("../states/Washington/output/Washington_Spring_2016.csv", gps)
        summer_2016 = process("../states/Washington/output/Washington_Summer_2016.csv", gps)
        
        spring_2017 = process("../states/Washington/output/Washington_Spring_2017.csv", gps)
        summer_2017 = process("../states/Washington/output/Washington_Summer_2017.csv", gps)
        
        spring_2018 = process("../states/Washington/output/Washington_Spring_2018.csv", gps)
        summer_2018 = process("../states/Washington/output/Washington_Summer_2018.csv", gps)
        
        spring_2019 = process("../states/Washington/output/Washington_Spring_2019.csv", gps)
        summer_2019 = process("../states/Washington/output/Washington_Summer_2019.csv", gps)
        
        #2015
        plotter_wa(spring_2015, summer_2015, "2015", shp)
        # 2016
        plotter_wa(spring_2016, summer_2016, "2016", shp)
        # 2017
        plotter_wa(spring_2017, summer_2017, "2017", shp)
        # 2018
        plotter_wa(spring_2018, summer_2018, "2018", shp)
        # 2019
        plotter_wa(spring_2019, summer_2019, "2019", shp)
    }
}
```

```{r, fig.width=18, fig.asp=.5}
state("California")
state("Washington")
state("Nevada")
```










